---
title: "portfolio 8"
output: html_document
date: "2023-04-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(ggplot2)
library(dplyr)
df <- read_excel("/Users/lindsaystall/Downloads/PIRUS_Public_May2020.xlsx")
df[df == -99 | df == -88] <- NA
```

```{r}
duration_counts <- table(df$Rad_duration)

duration_df <- data.frame(duration = names(duration_counts),
                          count = as.numeric(duration_counts))

duration_df$duration <- factor(duration_df$duration, levels = c("1", "2", "3"))

ggplot(duration_df, aes(x = duration, y = count, fill = duration)) +
  geom_bar(stat = "identity") +
  labs(x = "Radicalization Duration",
       y = "Count",
       title = "Distribution of Radicalization Duration") +
  scale_fill_manual(values = c("palevioletred", "paleturquoise", "thistle2")) +
  theme_minimal()
```
1 = Short (less than a year)
2 = Medium (between one and five years)
3 = Long (more than five years)

Most radicalization took 1-5 years before violent action, though many also took five+ years. I was honestly surprised the radicalizaton process prior to violence was so lengthy.


```{r}
df%>%
  count(Event_Influence1) %>%
  arrange(desc(n))
```

Of the terrorists who cited a specific event for precipitating their attacks, the top events were the Vietnam War, 9/11, and Arab Spring/Syrian Civil War. The vast majority did not cite a specific even as spurring on their attacks.


```{r}
ggplot(df, aes(x = factor(Beliefs_Trajectory))) +
  geom_bar(fill = "mediumpurple") +
  labs(title = "Beliefs Trajectory Distribution", x = "Beliefs Trajectory", y = "Count") +
  scale_x_discrete(labels = c("Gradual", "Key moments"))
```

In line with the previous analyses, it is not surprising that of the data we have (excluding NA) most people cited a gradual belief radicalization.


```{r}
ggplot(df, aes(x = factor(Behaviors_Trajectory))) +
  geom_bar(fill = "mediumseagreen") +
  labs(title = "Behavior Trajectory Distribution", x = "Behavior Trajectory", y = "Count") +
  scale_x_discrete(labels = c("Gradual", "Key moments"))
```

Perhaps surprisingly, of the data we have (excluding NA) most people cited a key moment in spurring on their behavioral radicalization.


```{r}
df_filtered <- df[!is.na(df$Radicalization_Sequence),]

counts <- table(df_filtered$Radicalization_Sequence)

df_counts <- data.frame(sequence = names(counts), count = as.numeric(counts))

df_counts$sequence <- ifelse(df_counts$sequence == "1", "Beliefs first", 
                             ifelse(df_counts$sequence == "2", "Behaviors first", "Simultaneously"))


ggplot(df_counts, aes(x="", y=count, fill=sequence)) + 
  geom_bar(stat="identity", width=1) + 
  coord_polar("y", start=0) +
  scale_fill_manual(values=c("palevioletred", "paleturquoise", "thistle2"), 
                    labels=c("Beliefs first", "Behaviors first", "Simultaneously")) +
  theme_void() +
  labs(title="Radicalization Sequence")
```


Very interesting!!!! For most terrorists in the dataset, radicalization of behaviors preceded radicalization of beliefs, whereby the individual began undertaking illegal ideologically-motivated behavior before espousing radical beliefs. This totally goes against my preconceived notions. Very cool.


```{r}
df_filtered <- subset(df, !is.na(Prison_Radicalize))

ggplot(df_filtered, aes(x = factor(Prison_Radicalize), fill = factor(Prison_Radicalize))) +
  geom_bar() +
  scale_fill_discrete(name = "Prison Radicalization", labels = c("No", "Yes (before)", "Yes (after)", "Yes (in prison)")) +
  labs(title = "Prison Radicalization") +
  xlab("Prison Radicalization") +
  ylab("Count") 
```

0 = No, individual had reached maximum level of radicalization (in both behaviors and beliefs) before prison
1 = Yes, but radicalization had begun (in either behaviors or belief) before prison
26
2 = Yes, but radicalization reached maximum level (in both behaviors and belief) after prison
3 = Yes, radicalization began and reached maximum level in prison


Most radicalization reached maximum level (in both behaviors and belief) after prison